---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# shinyCRUK

<!-- badges: start -->
<!-- badges: end -->

**shinyCRUK** provides Cancer Research UK-branded components for building Shiny applications that align with CRUK's Helix Design System and brand guidelines. 

## Features

### CRUK Branding
- Pre-styled components following CRUK brand guidelines
- Consistent typography using Progress Medium and Poppins fonts
- CRUK color palette and visual identity
- Responsive layouts for mobile and desktop

### Key Components
- **Navigation**: Branded nav bar with responsive logos and selectors
- **Layout**: Central column containers for optimal content width
- **Fonts**: Page titles and headings with CRUK styling
- **Value Boxes**: Styled value boxes
- **Footers**: Complete footer with legal information and contact details
- **Utilities**: Rounding functions per CI guidelines

## Documentation

Comprehensive function documentation is available on the [package website](https://crukorg.github.io/shinyCRUK/).

Each function includes:

- Detailed parameter descriptions
- Usage examples
- Accessibility guidelines


## Installation
Install the development version from GitHub:
```r
# Using remotes
remotes::install_github("CRUKorg/shinyCRUK")
```

For some people this might fail, so try this instead:
```r
# First create a github token
# It opens a browser window where you'll create a token on the github website
usethis::create_github_token()

# You will need to make sure you have configured CRUK SSO on this token

# Then run this function and paste the token when prompted
gitcreds::gitcreds_set() 

# Then run this to download the package
remotes::install_github("CRUKorg/shinyCRUK", auth_token = gitcreds::gitcreds_get()$password) 
```

You may then want to add your token to the GITHUB_PAT environment variable. This should allow
you to run `remotes::install_github` without having to set the `auth_token` each time:
```r
Sys.setenv(GITHUB_PAT = gitcreds::gitcreds_get()$password)
```

## Quick Start

We've built a working Shiny app to showcase most of the features in the package, with the folllowing code:
<details><summary>Show code</summary>
```{r, eval=FALSE, code=readLines("./inst/examples/showcase-app/app.R")}

```
</details>

## Core Components

### Theme and Layout
```r
# Apply CRUK theme
tags$head(crukTheme())

# Create central content column (900px max width)
centralColumn(
  # Your content here
)
```


### Navigation
Navbar styling handles desktop and mobile, with up to two selectors in the navbar.
```r
# Simple navigation
crukNavTitle(Title = "Early Diagnosis Data Hub")

# With selectors
crukNavTitle(
  Title = "Data Dashboard",
  selectors = 2,
  selector1 = selectInput("year", "Year:", 2020:2024),
  selector2 = selectInput("region", "Region:", c("UK", "England", "Scotland", "Wales", "Northern Ireland"))
)
```

### Typography
```r
# Page title with optional subheading
crukTitle(
  "Cancer Incidence Trends",
  "Age-standardised rates per 100,000 population"
)
```

### Logos
```r
# Stacked logo (preferred)
crukLogo(height = "50px")

# Wide logo (use sparingly)
crukLogoWide(height = "35px")
```

### Value Boxes
```r
crukValueBox(
  id = "metric1",
  title = "Total Cases",
  value = crukRounding(1234),
  icon = shiny::icon("chart-bar"),
  icon_colour = "#00007e",  # CRUK blue
  detail = "Year ending 2024"
)
```

### Styled last review
Use these for the top of pages so users know how up to date material is.  
```r
# Public content
lastReview("20 October 2025")

# Health professional content
lastReview("20 October 2025", tag = "health_professional")

# Internal content
lastReview("20 October 2025", tag = "internal")

# Accepts multiple date formats
lastReview("20/10/2025")              # UK format
lastReview(as.Date("2025-10-20"))     # Date object
lastReview(Sys.Date())                # Current date
```

### Footer
```r
# Full footer with contact information
crukFooter()

# Footer without contact section 
crukFooter(includeContact = FALSE)
```

### Number Formatting
```r
# Format numbers per CRUK Intelligence guidelines
crukRounding(1200)
# Returns: "around 1,200"

crukRounding(156789)
# Returns: "around 157,000"
```

## Design System Integration

This package aligns with CRUK's Helix Design System which is used throughout the new Contentful pages. It is built to rely on [bslib](https://rstudio.github.io/bslib/) components, particularly `bslib::page_navbar()` for navigation layouts.

## Requirements

- shiny
- htmltools
- bslib
- dplyr

## Contributing

This package is maintained by the CRUK Cancer Intelligence team. For questions, issues, or contributions, open an issue on GitHub or message Matt or Sam.
